Genetic draft and valley crossing

### 2018-07-26 ms/valley_crossing.tex
updated explanation of alpha simulations

The complicating factor is that, in an asexual population, the dynamics are governed almost entirely by what happens in the nose of the fitness distribution, and these dynamics can be highly stochastic. In previous approaches, these stochastic effects were smoothed due to the presence of recombination \citep{neher_shraiman_2011, neher_shraiman_2009}.
We therefore must focus primarily on simulation approaches, as analytical solutions for the behavior of a mutation in the nose of an asexual population are difficult to obtain.
We show that fitness valley crossing occurs at an overall lower rate in rapidly adapting populations.
However, this is consistent with the fact that \emph{all} forms of adaptation are slowed down.
In addition, we confirm that in rapidly adapting populations, adaptive fixations of alleles are more likely to be complex adaptations involving a fitness valley than they are in populations where genetic drift is the primary force shaping genetic variation at linked sites; essentially, genetic draft maintains the linkage that makes it possible to leap across fitness landscapes rather than adapt primarily by climbing to local peaks.
These observations add to the growing intuition that complex adaptations that involve evolution across fitness valleys may not be a surprising result of the evolutionary process, but rather an expected one.

@@

Our model features a population of $N$ haploid individuals in which an effectively infinite number of ``background" loci, with weak fitness effects, are currently segregating.
We focus on two loci, which are initially fixed for alleles $a$ and $b$.
These alleles mutate to $A$ and $B$ at rates $\mu_A$ and $\mu_B$, respectively (we neglect back mutations).
Genotypes $ab$, $Ab$, $aB$, and $AB$ have fitnesses $1$, $1-\delta_A$, $1-\delta_B$, and $1+s$ respectively, with $\delta \geq 0$ (intermediate genotypes are deleterious) and $1/N < s < 1$ (double mutant $AB$ is strongly beneficial).
For the remainder of our analysis, we will assume that $\mu_A = \mu_B = \mu$ and $\delta_A = \delta_B = \delta$.
We further assume that the large number of loci of weak effect contribute to a constant background fitness variance $\sigma^2$.
By Fisher's ``fundamental theorem" of natural selection \citep{Fisher:1930}, the rate $v$ at which the mean fitness advances due to selection is set to $\sigma^{2}$.
If $x$ is the background fitness of a lineage, then the distribution of background fitnesses $f(x)$ is assumed to be roughly Gaussian: $f(x) = \frac{1}{\sqrt{2\pi}\sigma} \exp (\frac{(x-\bar{x})^2}{2\sigma^2})$.
Given this background fitness variance, we are interested in the expected time $\mathbb{E}\left[ T\right]$ for the double mutant $AB$ (our so-called ``complex adaptation'') to arise and fix in the population.
We emphasize that $\sigma^2$ is only a \emph{background} fitness variance and does not include the effects of the focal loci.
For the most part, we assume that the focal loci are either at low enough frequency that they do not substantially affect the total fitness variance or that they are on the way to fixation and the population is destined to become roughly monomorphic (until new beneficial background mutations can be introduced).

In very large populations, enough double mutants are generated that the valley crossing time is dominated by the time required for a double mutant to sweep through the population.
At the other extreme, in small populations, the crossing time is dominated by the wait for a deleterious single mutant to fix: only thereafter does a successful double mutant appear.
We zero in on the intermediate case: stochastic tunneling, the most interesting and mathematically demanding mode of valley crossing.
We will review existing theory for this case, discuss the difficulty in the mathematical analysis due to genetic draft, and then present results for valley crossing times using forward-time simulations across a range of population sizes that include the sequential fixation, tunneling, and deterministic fixation cases.

@@

Most of the time, a bubble will simply arise and go extinct.
If mutation and fixation are sufficiently rare events, then fixation of the double mutant can be modeled as a Poisson process: with probability $1-e^{-\mu \pfix W}$, the bubble with total size $W$ will give rise to a successful double mutant.
Computing the expected value of this quantity over all bubble sizes, $\phi = \left< 1-e^{\mu \pfix W} \right>$, gives the probability that some bubble will lead to a successful valley crossing.
One difficulty in computing the rate of tunneling in the high $N\sigma$ case (where genetic draft is significant) is that the expected bubble size and fixation probability will both turn out to depend on the background fitness of an individual.
That is to say, it is not possible to simply compute $\pfix$ and the weight distribution by themselves, but rather one must convolute both of them over the background fitness.
Individuals whose background fitness is near the nose of the fitness distribution will tend to give rise to longer-lived lineages, and if those individuals give rise to the complex adaptation, it will therefore likewise be more likely to fix.
Thus, $W$ will depend on the background fitness $x$ and the valley depth $\delta$, and $p_{\mathrm fix}$ will depend on $x$ and the peak height $s$.
These will have to be integrated over the fitness distribution $f(x)$ in order to obtain the total crossing probability $\Phi(s,\delta) = \int_{-\infty}^\infty \phi(x,s,\delta) f(x) dx$.
The expected valley crossing time then becomes $1/N\mu\Phi$.

We begin by considering the dynamics of $\pfix(x,s)$ before moving on to consider the distribution of bubble sizes.
We focus on the case where background mutations are frequent but of weak effect--that is, the mean effect of background mutations $\bar{\epsilon}$ is small compared to $\sigma$, but background mutations occur at a high rate $U_b$.
\iffalse
In this case, the exact values of $U_b$ and $\bar{\epsilon}$ turn out not to matter: what is relevant is the product $D = U_b \bar{\epsilon}^2/2$.
\fi
\citet{hallatschek_2011} and \citet{good_2014} showed that the fixation probability in this regime obeys the differential equation 
\begin{equation}
v\partial_x p(x,s) = x p(x,s) + \frac{U_b\bar{\epsilon}^2}{2}\partial_x^2 p(x,s) - \frac{p(x,s)^2}{2},
\end{equation}
in which we have further assumed that $v \approx \sigma^2$, which is tantamount to claiming that mutational effects are less important than selection in determining the advance of the mean fitness.
For a neutral mutation ($s = 0$), the fixation probability is effectively zero at low $x$ but increases exponentially up until $x_c$, at which it becomes roughly linear in $x$.
In this regime, $x_c \approx 2\sigma^4/U_b\bar{\epsilon}^2$: when $\sigma$ is large, only mutations in the nose stand any chance of fixing unless they can be dragged to higher fitness by a background mutations (large $D$).
The effect of introducing a nonzero $s$ is essentially to shift the value of $x$ and $x_c$: for large negative $s$, $x_c$ may be far enough in the nose that it is unlikely any individuals will actually be present, and for large positive $s$, even unfit individuals can give rise to a lineage that is destined to fix.

In order to obtain the total valley crossing probability $\Phi$, the only remaining element to compute is the distribution of mutant bubble sizes $W$ as function of their background fitness $x$.
Since the tunneling probability for a lineage with background fitness $x$ is given by $\phi = \left< 1 - e^{-\mu \pfix W} \right>$, we do not need to compute the full distribution of $W$ but rather only its Laplace transform: $\mathcal{L}\left[ p(w) \right] = \int_0^\infty e^{-zw} p(w) dw$, with $z = \mu \pfix$.
\citet{neher_shraiman_2011} showed that $\phi$ follows the equation
\begin{equation}
\partial_t \phi = z + (z + x - \sigma^2 t + \delta)\phi^2 - (1 + x - \sigma^2 t + \delta)\phi^2,
\end{equation}
but their analysis also included a recombination term that made it possible to pull out the integrated Laplace transform $\int_{-\infty}^\infty p(x) \phi(x) dx$ without the need to actually solve this equation.
They also avoided the need to explicitly send time to infinity, which is necessary to incorporate all possible bubble sizes and lifetimes.
Unfortunately this method is not available to us, and the nose dominance of the bubble size distribution, combined with the $x$ dependence of $z$ (via $z = \mu \pfix(x)$), lead to difficulty in deriving a simple closed form expression for $\Phi$.
Therefore, we focus on simulation methods to study the dynamics of valley crossing.

@@

Simulations are allowed to proceed for an equilibration time of $N/10$ generations to allow sufficient genetic diversity to be introduced.
During this time, the mutation rate $\mu$ for the focal loci is set to zero.
After equilibration, we set a per-site mutation rate $\mu$ for the focal loci (allowing forward and backward mutations) with single mutant fitness disadvantage $\delta$ and double mutant advantage $s$.
When the double mutant reaches frequency $0.5$, we consider it to have fixed.
We then set the focal loci allele frequencies back to zero, randomize the selection coefficients for the genetic background by drawing anew from an exponential distribution, and allow the population to equilibrate for another $N/10$ generations.
This ensures that independent valley crossing trials are independent despite occurring in the same simulation run.

We partitioned our simulations as follows.
We first studied the effect of increasing $\sigma$ in the sequential fixation, stochastic tunneling, semi-deterministic tunneling, and sequential fixation regimes \citep{weissman_2009}.
\jcomment{commented out some text here; put back into the results if the pattern described is evident there}
% In the case of sequential fixation, the valley crossing time is dominated by the time until a deleterious mutant that is destined to fix arises: once it does, it does so in order $T_{\mathrm{MRCA}}$ generations, which is of order $N$ for low $N\sigma$ and scales weakly with $N$ for higher $N\sigma$ \citep{neher_hallatschek_2013}.
For small populations, sequential fixation applies when $N \ll \min(1/\delta, 1/\sqrt{\mu s})$, so that a successful mutant is unlikely to arise until the intermediate has fixed, but the intermediate's fitness disadvantage is too weak for selection to inhibit it.
Tunneling becomes relevant when $N$ is too large for sequential fixation, i.e., the tunneling probability becomes more significant than the probability that the intermediate will fix: this mandates $N \gtrsim 1/\sqrt{\mu s}$.
It also requires that $N \ll 1/\mu$, meaning that only one mutant lineage is likely to be extant at any given time and, if lineages do co-occur, they are unlikely to interfere with each other or substantially affect the mean fitness.
By increasing $N$ far above the $1/\mu$ limit, multiple deleterious mutants are likely to occur every generation, reaching mutation-selection balance at low frequency and providing many opportunities for a lucky double mutant to appear; this is semi-deterministic tunneling.
At still higher population sizes ($N \gtrsim s/\mu^2$), multiple double mutants are likely to appear every generation, and their dynamics can be modeled as deterministic.

@@

\begin{figure}
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/det_tunnel.pdf}
\end{subfigure}
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/det_fix.pdf}
\end{subfigure}
\caption{Valley crossing times in the semi-deterministic tunneling (left) and deterministic fixation (right) regimes as a function of $\sigma$. In both figures, $s = 10^{-2}$ and $\delta = 10^{-3}$: on the left, $N = 10^5$ and $\mu = 10^{-4}$, and on the right, $N = 10^6$ and $\mu = 10^{-3}$. Orange lines are averages, boxes are interquartile ranges, and end caps are overall ranges. Each box represents about $100$ simulation runs.}
\label{fig:deterministic}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.4\textwidth]{Figures/compare_ratios.pdf}
\caption{Ratio of $\alpha$ values at high and low $\sigma$. Genetic draft is helpful for crossing deep valleys in the sequential fixation and tunneling regimes. But when semi-deterministic tunneling becomes more important than stochastic tunneling, it can slow down the crossing rate.}
\label{fig:alpha_ratio}
\end{figure}

@@

We can get a clearer picture of how draft affects valley crossing times relative to single beneficial sweeps by considering the ratio $\alpha(10^{-2})/\alpha(10^{-6})$, which is a measure of the extent to which rapid adaptation speeds up valley crossing relative to sweeping (lower values indicate that rapid adaptation favors valley crossing).
The ratio can be seen in Figure \ref{fig:alpha_ratio}.
Overall, we find that raising $\sigma$ increases the extent to which valley crossing is favored for lower population sizes: for this combination of parameters, the condition is $N < 10^{5}$.
In this regime, raising $\delta$ also increases the degree to which valley crossing is favored under draft, which means that the deeper the valley, the more that draft increases the likelihood of valley crossing relative to a single beneficial mutation.
For $N > 10^5$ in the low $\sigma$ case, the product $N\mu > 1$: tunneling becomes semi-deterministic and, on average, more than one mutant lineage is produced every generation.
These mutants then give rise to double mutants at a low but constant rate that is not sensitive to the valley depth (Figure \ref{fig:tau-ratios} and equation 35 in \citealp{weissman_2009}) and that is much faster than in the stochastic tunneling case.
In contrast, $N > 10^{5}$ for high $\sigma$ does not produce the same order of magnitude increase in valley crossing rates, which is likely due to no analogous qualitative change in valley crossing dynamics.
Thus, for large enough population sizes, draft does not significantly improve the valley crossing rate relative to the rate of single beneficial sweeps.
\iffalse
\jcomment{Double check my explanation here :-)}
At high $\sigma$, the number of intermediate mutants is likely to fluctuate substantially \citep{Cvijovic:Good:2018}, but the average behavior is more or less the same.
This relation more or less obtains in a rapidly adapting population as well, with substantial fluctuations, so $\pfix$ becomes more important in determining the valley crossing rate: since this scales more weakly with $s$ under draft, the overall crossing rate is slowed.\jcomment{I still don't get what the point of mentioning this is here; how does this connection to the behavior of the ratio of $\alpha$s?}
\fi

>>>>

The complicating factor is that, in an asexual population, the dynamics are governed almost entirely by what happens in the nose of the fitness distribution, and these dynamics can be highly stochastic. In previous approaches, these stochastic effects were smoothed due to the presence of recombination \citep{neher_shraiman_2011, neher_shraiman_2009}.
We therefore must focus primarily on simulation approaches, as analytical solutions for the behavior of a mutation in the nose of an asexual population are difficult to obtain.
We show that fitness valley crossing occurs at an overall lower rate in rapidly adapting populations.
However, this is consistent with the fact that \emph{all} forms of adaptation are slowed down due to clonal interference and genetic background effects.
In addition, we confirm that in rapidly adapting populations, adaptive fixations of alleles are more likely to be complex adaptations involving a fitness valley than they are in populations where genetic drift is the primary force shaping genetic variation at linked sites; essentially, genetic draft maintains the linkage that makes it possible to leap across fitness landscapes rather than adapt primarily by climbing to local peaks.
These observations add to the growing intuition that complex adaptations that involve evolution across fitness valleys may not be a surprising result of the evolutionary process, but rather an expected one.

@@

Our model features a population of $N$ haploid individuals in which an effectively infinite number of ``background" loci, with weak fitness effects, are currently segregating.
We focus on two loci, which are initially fixed for alleles $a$ and $b$.
These alleles mutate to $A$ and $B$ at rates $\mu_A$ and $\mu_B$, respectively (we neglect back mutations).
Genotypes $ab$, $Ab$, $aB$, and $AB$ have fitnesses $1$, $1-\delta_A$, $1-\delta_B$, and $1+s$ respectively, with $\delta \geq 0$ (intermediate genotypes are deleterious) and $1/N < s < 1$ (double mutant $AB$ is strongly beneficial); see \ref{fig:landscape}.
For the remainder of our analysis, we will assume that $\mu_A = \mu_B = \mu$ and $\delta_A = \delta_B = \delta$.
We further assume that the large number of loci of weak effect contribute to a constant background fitness variance $\sigma^2$.
By Fisher's ``fundamental theorem" of natural selection \citep{Fisher:1930}, the rate $v$ at which the mean fitness advances due to selection is set to $\sigma^{2}$.
If $x$ is the background fitness of a lineage, then the distribution of background fitnesses $f(x)$ is assumed to be roughly Gaussian: $f(x) = \frac{1}{\sqrt{2\pi}\sigma} \exp (\frac{(x-\bar{x})^2}{2\sigma^2})$.
Given this background fitness variance, we are interested in the expected time $\mathbb{E}\left[ T\right]$ for the double mutant $AB$ (our so-called ``complex adaptation'') to arise and fix in the population.
We emphasize that $\sigma^2$ is only a \emph{background} fitness variance and does not include the effects of the focal loci.
For the most part, we assume that the focal loci are either at low enough frequency that they do not substantially affect the total fitness variance or that they are on the way to fixation and the population is destined to become roughly monomorphic (until new beneficial background mutations can be introduced).

In very large populations, enough double mutants are generated that the valley crossing time is dominated by the time required for a double mutant to sweep through the population.
At the other extreme, in small populations, the crossing time is dominated by the wait for a deleterious single mutant to fix: only thereafter does a successful double mutant appear.
We zero in on the intermediate case: stochastic tunneling, the most interesting and mathematically demanding mode of valley crossing.
We will review existing theory for this case, discuss the difficulty in the mathematical analysis due to genetic draft, and then present results for valley crossing times using forward-time simulations across a range of population sizes that include the sequential fixation, stochastic tunneling, and semi-deterministic tunneling cases.

@@

Most of the time, a bubble will simply arise and go extinct.
If mutation and fixation are sufficiently rare events, then fixation of the double mutant can be modeled as a Poisson process: with probability $1-e^{-\mu \pfix W}$, the bubble with total size $W$ will give rise to a successful double mutant.
Computing the expected value of this quantity over all bubble sizes, $\phi = \left< 1-e^{\mu \pfix W} \right>$, gives the probability that some bubble will lead to a successful valley crossing.
One difficulty in computing the rate of tunneling in the high $N\sigma$ case (where genetic draft is significant) is that the expected bubble size and fixation probability will both turn out to depend on the background fitness of an individual.
That is to say, it is not possible to simply compute $\pfix$ and the weight distribution by themselves, but rather one must convolute both of them over the background fitness.
Individuals whose background fitness is near the nose of the fitness distribution will tend to give rise to longer-lived lineages, and if those individuals give rise to the complex adaptation, it will therefore likewise be more likely to fix.
Thus, $W$ will depend on the background fitness $x$ and the valley depth $\delta$, and $p_{\mathrm fix}$ will depend on $x$ and the peak height $s$.
These will have to be integrated over the fitness distribution $f(x)$ in order to obtain the total crossing probability $\Phi(s,\delta) = \int_{-\infty}^\infty \phi(x,s,\delta) f(x) dx$.
The expected valley crossing time then becomes $1/N\mu\Phi$.
An additional difficulty is that technically, $\pfix$ will depend not on the fitness of the initial lineage in a bubble but on the fitness of the specific background on which the double mutant arose.
These will not necessarily be the same; for example, individuals within the single mutant lineage may accumulate additional background mutations.
Assuming that this does \emph{not} happen, that is, assuming that $\pfix$ depends on the fitness $x$ of the founding single mutant lineage, is tantamount to assuming that bubbles are too short lived to accumulate background mutations that significantly affect their fitness.

We begin by considering the dynamics of $\pfix(x,s)$ before moving on to consider the distribution of bubble sizes.
We focus on the case where background mutations are frequent but of weak effect--that is, the mean effect of background mutations $\bar{\epsilon}$ is small compared to $\sigma$, but background mutations occur at a high rate $U_b$.
\iffalse
In this case, the exact values of $U_b$ and $\bar{\epsilon}$ turn out not to matter: what is relevant is the product $D = U_b \bar{\epsilon}^2/2$.
\fi
\citet{hallatschek_2011} and \citet{good_2014} showed that the fixation probability in this regime obeys the differential equation 
\begin{equation}
v\partial_x p(x,s) = x p(x,s) + \frac{U_b\bar{\epsilon}^2}{2}\partial_x^2 p(x,s) - \frac{p(x,s)^2}{2},
\end{equation}
in which we have further assumed that $v \approx \sigma^2$, which is tantamount to claiming that mutational effects are less important than selection in determining the advance of the mean fitness.
For a neutral mutation ($s = 0$), the fixation probability is effectively zero at low $x$ but increases exponentially up until $x_c$, at which it becomes roughly linear in $x$.
In this regime, $x_c \approx 2\sigma^4/U_b\bar{\epsilon}^2$: when $\sigma$ is large, only mutations in the nose stand any chance of fixing unless they can be dragged to higher fitness by background mutations.
The effect of introducing a nonzero $s$ is essentially to shift the value of $x$ and $x_c$: for large negative $s$, $x_c$ may be far enough in the nose that it is unlikely any individuals will actually be present, and for large positive $s$, even unfit individuals can give rise to a lineage that is destined to fix.

In order to obtain the total valley crossing probability $\Phi$, the only remaining element to compute is the distribution of mutant bubble sizes $W$ as function of their background fitness $x$.
Since the tunneling probability for a lineage with background fitness $x$ is given by $\phi = \left< 1 - e^{-\mu \pfix W} \right>$, we do not need to compute the full distribution of $W$ but rather only its Laplace transform: $\mathcal{L}\left[ p(w) \right] = \int_0^\infty e^{-zw} p(w) dw$, with $z = \mu \pfix$.
\citet{neher_shraiman_2011} showed that $\phi$ follows the equation
\begin{equation}
\partial_t \phi = z + (z + x - \sigma^2 t + \delta)\phi - (1 + x - \sigma^2 t + \delta)\phi^2,
\end{equation}
but their analysis also included a recombination term that made it possible to pull out the integrated Laplace transform $\int_{-\infty}^\infty p(x) \phi(x) dx$ without the need to actually solve this equation.
They also avoided the need to explicitly send time to infinity, which is necessary to incorporate all possible bubble sizes and lifetimes.
Unfortunately this method is not available to us, and the nose dominance of the bubble size distribution, combined with the $x$ dependence of $z$ (via $z = \mu \pfix(x)$), lead to difficulty in deriving a simple closed form expression for $\Phi$.
Therefore, we focus on simulation methods to study the dynamics of valley crossing.

@@

Simulations are allowed to proceed for an equilibration time of $N/10$ generations to allow sufficient genetic diversity to be introduced.
During this time, the mutation rate $\mu$ for the focal loci is set to zero.
After equilibration, we set the per-site mutation rate $\mu$ for the focal loci (allowing forward and backward mutations) with single mutant fitness disadvantage $\delta$ and double mutant advantage $s$.
When the double mutant reaches frequency $0.5$, we consider it to have fixed.
We then set the focal loci allele frequencies back to zero, randomize the selection coefficients for the genetic background by drawing anew from an exponential distribution, and allow the population to equilibrate for another $N/10$ generations.
This ensures that independent valley crossing trials are independent despite occurring in the same simulation run.

We partitioned our simulations as follows.
We first studied the effect of increasing $\sigma$ in the sequential fixation, stochastic tunneling, semi-deterministic tunneling, and deterministic fixation regimes \citep{weissman_2009}.
% In the case of sequential fixation, the valley crossing time is dominated by the time until a deleterious mutant that is destined to fix arises: once it does, it does so in order $T_{\mathrm{MRCA}}$ generations, which is of order $N$ for low $N\sigma$ and scales weakly with $N$ for higher $N\sigma$ \citep{neher_hallatschek_2013}.
For small populations, sequential fixation applies when $N \ll \min(1/\delta, 1/\sqrt{\mu s})$, so that a successful mutant is unlikely to arise until the intermediate has fixed, but the intermediate's fitness disadvantage is too weak for selection to inhibit it.
Tunneling becomes relevant when $N$ is too large for sequential fixation, i.e., the tunneling probability becomes more significant than the probability that the intermediate will fix: this mandates $N \gtrsim 1/\sqrt{\mu s}$.
It also requires that $N \ll 1/\mu$, meaning that only one mutant lineage is likely to be extant at any given time and, if lineages do co-occur, they are unlikely to interfere with each other or substantially affect the mean fitness.
By increasing $N$ far above the $1/\mu$ limit, multiple deleterious mutants are likely to occur every generation, reaching mutation-selection balance at low frequency and providing many opportunities for a lucky double mutant to appear; this is semi-deterministic tunneling.
At still higher population sizes ($N \gtrsim s/\mu^2$), multiple double mutants are likely to appear every generation, and their dynamics can be modeled as deterministic.

@@

\begin{figure}
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/det_tunnel.pdf}
\end{subfigure}
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/det_fix.pdf}
\end{subfigure}
\caption{Valley crossing times in the semi-deterministic tunneling regime as a function of $\sigma$. In both figures, $s = 10^{-2}$ and $\delta = 10^{-3}$: on the left, $N = 10^5$ and $\mu = 10^{-4}$, and on the right, $N = 10^6$ and $\mu = 10^{-3}$; the right figure has a high enough value of $N\mu^2$ that fixation is almost deterministic. Orange lines are averages, boxes are interquartile ranges, and end caps are overall ranges. Each box represents about $100$ simulation runs.}
\label{fig:deterministic}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.95\textwidth]{Figures/compare_ratios_multi.pdf}
\caption{Ratio of $\alpha$ values at high and low $\sigma$, for varied values of $\mu$. Genetic draft is helpful for crossing deep valleys in the sequential fixation and tunneling regimes, though at large population sizes it can slow down the crossing rate for deep valleys: at higher values of $\mu$ this limitation disappears.}
\label{fig:alpha_ratio}
\end{figure}

@@

We can get a clearer picture of how draft affects valley crossing times relative to single beneficial sweeps by considering the ratio $\alpha(10^{-2})/\alpha(10^{-6})$ (the arguments are values of $\sigma$), which is a measure of the extent to which rapid adaptation speeds up valley crossing relative to sweeping (lower values indicate that rapid adaptation favors valley crossing).
The ratio can be seen in Figure \ref{fig:alpha_ratio}.
Overall, we find that raising $\sigma$ increases the extent to which valley crossing is favored for lower population sizes: for this combination of parameters, the condition is $N < 10^{5}$.
In this regime, raising $\delta$ also increases the degree to which valley crossing is favored under draft, which means that the deeper the valley, the more that draft increases the likelihood of valley crossing relative to a single beneficial mutation.
The exception is at very high values of $N$ and $\delta$ and low values of $\mu$, where draft turns out again not to be helpful; see the lower right part of the leftmost plot.
In this regime, $\delta$ is on the order of $\sigma$ in our draft simulations, meaning that single mutant lineages are rapidly cut off by the advancing mean fitness and are likely almost never present in the nose of the fitness distribution.
Essentially, the population size is large enough that intermediate mutants can be ``seen" by selection, but the fitness detriment is severe enough compared to $\sigma$ that draft is no longer helpful in allowing mutant lineages to reach high frequency.
This problem disappears as $\mu$ increases (center and right plots) and the population is far beyond the semi-deterministic tunneling barrier.
